---
title: " Impact pathways and forecast of Integrated Soil Fertility Management in northern Ghana"
author: "Dorcas, Cory, Javier, Eike"
date: "2025-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("gtExtras")
#install.packages("svglite")
#library(gtExtras)
#library(svglite)

```

**Description**
Using probabilistic modeling approaches, we are modeling the farm level benefits of Integrated Soil Fertility Management (ISFM), a sustainable intensification practice. ISFM holds promise to improve soil fertility and increase productivity on the already exciting land for small-holder farmers in northern Ghana. ISFM is a set of the following components (below), starting with a few components as partial ISFM until complete ISFM (Vanlawe et al., 2010: https://doi.org/10.5367/0000000107911699)

**Improved seed**
As an entry point to ISFM, using improved and resilient seed has its own benefits and costs incurred as they are more expensive than the farmers seed. However farmers are aware that these improved seed yield more than their own seed. 

**Inorganic fertilizer**
Because soils are poor, they need to be replenished with additional nutrients. Using inorganic fertilizer such as NPK and urea on improved seed is the second layer in ISFM. This second layer of ISFM involves additional costs of buying the fertilizer and additional labor for fertilizer application.

**Organic fertilizer**
This includes using improved seed and the use of already available organic material such as crop residue and animal manure. However farmers are not always able to apply the amount of organic fertilizer needed for sufficient fertilization because the required amount is usually high.

**Fertilizer combination**
Because each type of fertilizer (organic or inorganic) has its unique characteristic and benefit, combining the two will have synergistic benefits that will be more beneficial than applying one alone on improved seed. Combining them also reduces the amount applied for each compared to when they are applied alone.

**Minimum tillage**
To avoid the compaction of soil with heavy machinery, tillage is recommended. However in the context of northern Ghana where zero tillage is almost impossible due to long drought and the nature of their soils, minimum tillage is recommended. Here farmers are also advised to avoid using tractors on their soils to tilt the land. 

**Complete ISFM**
When all the above components are fully adopted simultaneously. 

The 6 different practices are the decision options in a maize and soybean rotation system against the baseline of maize monoculture: 
1. Use of Improved seed (As the entry point to ISFM)
2. Improved seed and inorganic fertilizer 
3. Improved seed and organic fertilizer 
4. Improved seed and fertilizer combination (half organic and inorganic fertilizer)
5. Improved seed and minimum tillage
6. Improved seed, fertilizer combination and minimum tillage (Complete/full ISFM)

The model will answer the following question: 
**Should Agricultural Research and Development (AR4D) actors promote ISFM in northern Ghana? **

The impact pathways of ISFM incorporate the costs, the benefits and the risks of each ISFM component for a decade because estimations related to soil are mostly done for a decade so we used this as the basis for our time frame (Guibert, 1999). The impact pathway was translated into a mathematical model and run in montecarlo simulation to find the plausible impact of ISFM and the most economic and sustainable ISFM business model. 


# Montecarlo simulation for ISFM benefits and tradeoffs in northern Ghana 

Using the pre-built functions for the 6 ISFM decision options, we run an mc simulation  

```{r mcSimulation, warning=FALSE, size= "huge", message=FALSE}
# Source our model from the model script
source("functions/ISFM_components_function.R")

# Setting seed to ensure consistent results
#each time we run the entire simulation 
set.seed(233) 

ISFM_mc_simulation <- mcSimulation(as.estimate(ISFM_table), 
                              model_function = ISFM_components_function,
                              numberOfModelRuns = 10000,
                              functionSyntax = "plainNames")
```

Plotting the different outcomes of the 6 decision options to find the most sustainable and economically viable option of ISFM based on the sustainable intensification assessment framework indicators. The most sustainable and beneficial option will be the one that will be further assessed based on the different farmers typology later. We first arrange the montecarlo simulation data

```{r}
#Arranging the Monte Carlo simulation data first
library(tidyverse)
library(tidyr)
library(stringr)

# Extract values from the simulation output 
ISFM_impact_data <- ISFM_mc_simulation$y[1:35]

# Pivot the data to long format
ISFM_pivoted_outcome_data <- tidyr::pivot_longer(ISFM_impact_data,
                                                 cols = everything(), 
                                                 names_to = "Outcome with ISFM", 
                                                 values_to = "Absolute_difference")

# Extract outcome type
ISFM_pivoted_outcome_data["Outcome"] <- 
  str_extract(ISFM_pivoted_outcome_data$`Outcome with ISFM`, 
  "productivity|income|environmental|social|human")

# Extract ISFM practice type by removing the identified outcome from the original column name
ISFM_pivoted_outcome_data["Practices"] <-  
  str_remove(ISFM_pivoted_outcome_data$`Outcome with ISFM`, 
  "productivity|income|environmental|social|human")

```


To find the summary of the ISFM outcome data 

```{r}
library(dplyr)

# Calculate quantiles for the distribution
ISFM_outcome_bounds <- ISFM_pivoted_outcome_data %>%
  group_by(Outcome, Practices) %>%
  summarise(
    lower_bound = quantile(Absolute_difference, 0.05, na.rm = TRUE),  # 5th percentile for lower bound
    upper_bound = quantile(Absolute_difference, 0.95, na.rm = TRUE),  # 95th percentile for upper bound
    min_value = min(Absolute_difference, na.rm = TRUE),               # Minimum value of the distribution
    max_value = max(Absolute_difference, na.rm = TRUE),               # Maximum value of the distribution
    mean_value = mean(Absolute_difference, na.rm = TRUE),             # Mean of the distribution
    sd_value = sd(Absolute_difference, na.rm = TRUE)                  # Standard deviation of the distribution
  ) %>%
  ungroup()

ISFM_outcome_bounds

```


We generate a radar (spider) plot that visualizes trade-offs between ISFM practices across different sustainable intensification outcomes (productivity, economic, environmental, social and human). These outcomes are compared to the statusquo.

```{r}
#install.packages("ggradar", repos = "https://cloud.r-project.org")
#install.packages("scales")
#install.packages("ggplot2")

library(ggradar)
library(ggplot2)
library(scales)

# We use means to summarize the outcome data for each practice
ISFM_summary_data <- ISFM_pivoted_outcome_data %>%
  group_by(Practices, Outcome) %>%
  summarize(Absolute_difference = mean(Absolute_difference, na.rm = TRUE), .groups = "drop")

#Transforming the data to wide format to easily visualize into a spider plot
ISFM_radar_data <- ISFM_summary_data %>%
  pivot_wider(names_from = Outcome, values_from = Absolute_difference, values_fill = list(Absolute_difference = 0))
ISFM_radar_data[, -1] <- lapply(ISFM_radar_data[, -1], as.numeric)

# Because we are dealing for large numbers and outcomes have different units, 
#we normalize the outcomes values between 0 and 1 for visualization

ISFM_radar_data[, -1] <- apply(ISFM_radar_data[, -1], 2, function(x) rescale(x, to = c(0, 1)))

ISFM_radar_data <- ISFM_radar_data %>%
  rename(group = Practices)

ISFM_radar_plot <- ggradar(ISFM_radar_data,
        grid.min = 0, grid.mid = 0.5, grid.max = 1,  
        values.radar = c("0", "0.5", "1"),
        group.colours = rainbow(nrow(ISFM_radar_data)),
        legend.position = "bottom") +
  ggtitle("Trade-offs between ISFM practices across sustainable intensification outcomes") +
  theme(
    text = element_text(size = 0.1),
    plot.title = element_text(size = 15, face = "bold"),
    legend.text = element_text(size = 8)  
  ) + theme(aspect.ratio = 0.5)

ISFM_radar_plot

ggsave("figures/ISFM_radar_plot.png", width = 8, height = 5, dpi = 300)

```

To assess the stable and most profitable ISFM component under uncertainty we did a stochastic dominance test. 
Stochastic dominance is a sophisticated comparison of distributions with the dominating distribution being superior in terms of mean, variance, skewness and kurtosis. Stochastic dominance tests help to rank different distributions i.e the superiority of one probability distribution over another https://cran.r-project.org/web/packages/stodom/stodom.pdf. We will only limit to the first oder of dominance where farmers utility is based on maximizing the benefits (more to less).
In comparision to the Decision Support package, the stochastic dominance method does not directly generate distributions based on uncertainty, but instead, it helps in comparing existing distributions of outcomes to see which one is preferred, based on certain dominance rules. It provides a way to determine whether one option (or distribution) is "better" than another without the need for specific risk preferences or utility functions. We will only test first order stochastic dominance since the scope of our study is only on utility and not on risk analysis.

```{r}

# We compute the cumulative probabilities
ISFM_cdf_data <- ISFM_pivoted_outcome_data %>%
  filter(!is.na(Absolute_difference)) %>%
  group_by(Practices, Outcome) %>%
  arrange(Absolute_difference) %>%
  mutate(cdf = row_number() / n())

# We first add the units label and reorder the outcomes and practices
ISFM_cdf_data <- ISFM_cdf_data %>%
  mutate(
    Outcome_label = case_when(
      Outcome == "productivity" ~ "Productivity (kg/acre)",
      Outcome == "income" ~ "Income (Ghana cedis/acre)",
      Outcome == "environmental" ~ "Environmental benefits (Ghana cedis/acre)",
      Outcome == "social" ~ "Social benefits (Ghana cedis/acre)",
      Outcome == "human" ~ "Human benefits (Ghana cedis/acre)",
      TRUE ~ Outcome
    ),
    Outcome_label = factor(Outcome_label, levels = c(
      "Productivity (kg/acre)",
      "Income (Ghana cedis/acre)",
      "Environmental benefits (Ghana cedis/acre)",
      "Social benefits (Ghana cedis/acre)",
      "Human benefits (Ghana cedis/acre)"
    )),
    Practices = factor(Practices, levels = c(
      "statusquo_", 
      "improved_seed_", 
      "mineral_fertilizer_", 
      "organic_fertilizer_",
      "fertilizer_combination_",
      "tillage_", 
      "complete_ISFM_"
    ))
  )

# Plotting the distribution
cdf_ISFM_plot <-
  ggplot(ISFM_cdf_data, aes(x = Absolute_difference, y = cdf, color = Practices)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~Outcome_label, scales = "free", ncol = 1, drop = FALSE) +
  labs(
    title = " ",
    x = "Absolute Difference",
    y = "Cumulative Probability",
    color = "ISFM Practices"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    text = element_text(size = 12),
    strip.text = element_text(size = 11, face = "bold"),  
    panel.spacing = unit(1.5, "lines")
  )

cdf_ISFM_plot

ggsave("figures/cdf_ISFM_plot.png", width = 8, height = 10, dpi = 300)

```


Further, because farmers are not heterogeneous, they differ based on resources available to them, we identified important resources that could impact ISFM outcomes. We account for them as typology and assess how the most beneficial ISFM is affected by these vital resources. 

```{r typology_graph, warning=FALSE, message=FALSE}

factors <- data.frame(
  Ressources = c("Land", "Agricultural inputs/Financial ressources","Technical skills/knowledge",
                 "Labor", "Market Linkages", 
                 "Extension services",
            "Utilities (electricity, internet)", "Time"),
  Strength = c(5, 5, 4, 4, 3, 3, 0, 3)
)

ggplot(factors, aes(x=Strength, y=reorder(Ressources, Strength))) +
  geom_point(size=5, color="darkgreen") +
  theme_minimal() +
  labs(x="Strength", y="Ressources") +
  geom_segment(aes(xend=0, yend=Ressources), color="lightgray")

ggsave("figures/typology_graph.png", width = 8, height = 5, dpi = 300)


```

Based on these identified typologies we present the economic return of full ISFM components (complete adoption) which  encompasses the income based on available resources and gender. We represent this as NPVs and cumulative cashflows depending on the identified vital resources for ISFM: land, agricultural inputs, knowledge, and labor. 

We first find the summary of NPVs based on gender and resources

```{r}
range_NPV <- (ISFM_mc_simulation$y)[36:45]

library(dplyr)

summary_ranges <- range_NPV %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarise(
    `5th Percentile` = quantile(Value, probs = 0.05, na.rm = TRUE),
    `95th Percentile` = quantile(Value, probs = 0.95, na.rm = TRUE),
    Min = min(Value, na.rm = TRUE),
    Median = median(Value, na.rm = TRUE),
    Max = max(Value, na.rm = TRUE)
  )

```

Plotting the income distribution based on gender and vital resources.

```{r}

library(RColorBrewer)

# Define the ordered list of variable names
ordered_vars <- c(
  "NPV_men_ISFM_income",
  "NPV_women_ISFM_income",
  "NPV_men_ISFM_income_land_based",
  "NPV_women_ISFM_income_land_based", 
  "NPV_men_ISFM_income_inputs_based",
  "NPV_women_ISFM_income_inputs_based", 
  "NPV_men_ISFM_income_knowledge_based",
  "NPV_women_ISFM_income_knowledge_based",
  "NPV_men_ISFM_income_labor_based",
  "NPV_women_ISFM_income_labor_based"
)

# Extract the corresponding data from the mcSimulation object
ISFM_data_violin <- as.data.frame(ISFM_mc_simulation$y)[, ordered_vars]

# Reshape the data to long format
ISFM_data_long <- pivot_longer(ISFM_data_violin, 
                          cols = everything(), 
                          names_to = "Scenario", 
                          values_to = "NPV")

# Ensure the variable order is preserved
ISFM_data_long$Scenario <- factor(ISFM_data_long$Scenario, levels = ordered_vars)

# Define a color palette with 10 distinct colors
palette <- brewer.pal(10, "Paired")

# Create the violin plot
ggplot(ISFM_data_long, aes(x = Scenario, y = NPV, fill = Scenario)) +
  geom_violin(trim = FALSE, scale = "width", color = "black") +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white") +
  scale_fill_manual(values = palette) +
  theme_minimal(base_size = 11) +
  labs(
    x = "Scenario",
    y = "Net Present Value (Ghana Cedis)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

ggsave("figures/NPV_ISFM_gender_resource.png", width = 8, height = 5, dpi = 300)

```

Cashflow analysis and plotting based on gender and farming resources

```{r}

ISFM_cashflow_men <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "men_cashflow_ISFM",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_cashflow_men


ISFM_cashflow_women <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "women_cashflow_ISFM",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_cashflow_women


ISFM_land_cashflow_men <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "men_cashflow_ISFM_land_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_land_cashflow_men

ISFM_land_cashflow_women <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "women_cashflow_ISFM_land_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_land_cashflow_women


ISFM_inputs_cashflow_men <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "men_cashflow_ISFM_inputs_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_inputs_cashflow_men

ISFM_inputs_cashflow_women <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "women_cashflow_ISFM_inputs_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_inputs_cashflow_women


ISFM_knowledge_cashflow_men <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "men_cashflow_ISFM_knowledge_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_knowledge_cashflow_men

ISFM_knowledge_cashflow_women <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "women_cashflow_ISFM_knowledge_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_knowledge_cashflow_women


ISFM_labor_cashflow_men <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "men_cashflow_ISFM_labor_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_labor_cashflow_men

ISFM_labor_cashflow_women <- plot_cashflow(mcSimulation_object = ISFM_mc_simulation,
              cashflow_var_name = "women_cashflow_ISFM_labor_based",
                          x_axis_name = "Years of intervention",
                          y_axis_name = "Cashflow in Ghana Cedis",
                          color_25_75 = "grey",
                          color_5_95 = "yellow",
                          color_median= "red",
                          base= 10 ) + ggplot2::scale_y_continuous(limits = c(0, 1000000))

ISFM_labor_cashflow_women

```


**Projection to Latent Structures (PLS)**

We use Projection to Latent Structures (PLS) model to assess the correlation strength and direction for model variables and outcome variables. The Partial Least Squares is fitted with the orthogonal scores algorithm with `pls::plsr`

```{r}
#To get the names of outcomes needed to run some posthoc test 
names(ISFM_mc_simulation$y)[36:45]
```

Here we only need to use the variables related to complete ISFM since it emerged to be the best ISFM. We remove variables for other ISFM so the correlation does not read the other ISFM components

```{r}

pls_ISFM_input_variables <-  ISFM_table %>%
  filter(!variable %in% c("decay_rate", "traditional_maize_seed_price", "maize_yield_statusquo",
                         "maize_yield_component1","maize_yield_component2",
                         "maize_yield_component3","maize_yield_component4",
                         "maize_yield_component5",  "soybean_yield_component1",
                        "soybean_yield_component2", "soybean_yield_component3", 
                        "soybean_yield_component4","soybean_yield_component5",
                         "nutrient_partial_balance_value_organic_fertilizer",     
                        "nutrient_partial_balance_value_mineral_fertilizer"))


# "NPV_women_ISFM" all available ressources
pls_women <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_women_ISFM_income",
                               ncomp = 1)

plot_pls_women <- plot_pls(plsrResults = pls_women, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)

# "NPV_women_ISFM based on land
pls_women_land <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_women_ISFM_income_land_based",
                               ncomp = 1)

plot_pls_women_land <- plot_pls(plsrResults = pls_women_land, 
                            input_table =pls_ISFM_input_variables , 
                            threshold = 1)

# "NPV_women_ISFM based on inputs
pls_women_inputs <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_women_ISFM_income_inputs_based",
                               ncomp = 1)

plot_pls_women_inputs <- plot_pls(plsrResults = pls_women_inputs, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)


# "NPV_women_ISFM based on knowledge
pls_women_knowledge <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_women_ISFM_income_knowledge_based",
                               ncomp = 1)

plot_pls_women_knowledge <- plot_pls(plsrResults = pls_women_knowledge, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)


# "NPV_women_ISFM based on labor
pls_women_labor <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_women_ISFM_income_labor_based",
                               ncomp = 1)

plot_pls_women_labor <- plot_pls(plsrResults = pls_women_labor, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)



# "NPV_men_ISFM" all available ressources
pls_men <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_men_ISFM_income",
                               ncomp = 1)

plot_pls_men <- plot_pls(plsrResults = pls_men, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)

# "NPV_men_ISFM based on land
pls_men_land <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_men_ISFM_income_land_based",
                               ncomp = 1)

plot_pls_men_land <- plot_pls(plsrResults = pls_men_land, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)

# "NPV_men_ISFM based on inputs
pls_men_inputs <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_men_ISFM_income_inputs_based",
                               ncomp = 1)

plot_pls_men_inputs <- plot_pls(plsrResults = pls_men_inputs, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)


# "NPV_men_ISFM based on knowledge
pls_men_knowledge <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_men_ISFM_income_knowledge_based",
                               ncomp = 1)

plot_pls_men_knowledge <- plot_pls(plsrResults = pls_men_knowledge, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)


# "NPV_women_ISFM based on labor
pls_men_labor <- plsr.mcSimulation(object = ISFM_mc_simulation,
                               resultName = "NPV_men_ISFM_income_labor_based",
                               ncomp = 1)

plot_pls_men_labor <- plot_pls(plsrResults = pls_men_labor, 
                            input_table = pls_ISFM_input_variables, 
                            threshold = 1)

```

Plotting all the PLS together

```{r}

# Adding titles to each plot
plot_pls_men             <- plot_pls_men + ggtitle("All Resources (Men)")
plot_pls_women           <- plot_pls_women + ggtitle("All Resources (Women)")

plot_pls_men_land        <- plot_pls_men_land + ggtitle("Land-based (Men)")
plot_pls_women_land      <- plot_pls_women_land + ggtitle("Land-based (Women)")

plot_pls_men_inputs      <- plot_pls_men_inputs + ggtitle("Inputs-based (Men)")
plot_pls_women_inputs    <- plot_pls_women_inputs + ggtitle("Inputs-based (Women)")

plot_pls_men_knowledge   <- plot_pls_men_knowledge + ggtitle("Knowledge-based (Men)")
plot_pls_women_knowledge <- plot_pls_women_knowledge + ggtitle("Knowledge-based (Women)")

plot_pls_men_labor       <- plot_pls_men_labor + ggtitle("Labor-based (Men)")
plot_pls_women_labor     <- plot_pls_women_labor + ggtitle("Labor-based (Women)")

# Combine plots by resource category
pls_by_resource_grid <- 
  (plot_pls_men + plot_pls_women) /
  (plot_pls_men_land + plot_pls_women_land) /
  (plot_pls_men_inputs + plot_pls_women_inputs) /
  (plot_pls_men_knowledge + plot_pls_women_knowledge) /
  (plot_pls_men_labor + plot_pls_women_labor) +
  plot_layout(guides = "collect") +
  plot_annotation(
   
    theme = theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12)
    )
  )

# Saving the final arranged pls plot
pls_by_resource_grid

ggsave("figures/ISFM_PLS_by_Resource_and_Gender.png",
       plot = pls_by_resource_grid,
       width = 15, height = 10, dpi = 300)


```


**Value of information **
Expected Value of Perfect Information (EVPI) 
This analysis is crucial since it informs the decision-maker how much he/she should be willing to pay if given perfect information, to eliminate uncertainties and avoid opportunity loss.

```{r}
evpi_mcresults_table <- data.frame(ISFM_mc_simulation$x,
                          ISFM_mc_simulation$y[36:45])

evpi<- multi_EVPI(mc= evpi_mcresults_table,
                  first_out_var = "NPV_women_ISFM_income", write_table = FALSE, outfolder = ./results/evpi_results)

```

Plotting EVPI 

```{r}

# Plot all EVPI for the decision variables 
women_evpi <- plot_evpi(evpi,
          decision_vars = "NPV_women_ISFM_income",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

men_evpi <- plot_evpi(evpi,
          decision_vars = "NPV_men_ISFM_income",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

women_evpi_land <- plot_evpi(evpi,
          decision_vars = "NPV_women_ISFM_income_land_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

men_evpi_land <- plot_evpi(evpi,
          decision_vars = "NPV_men_ISFM_income_land_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

women_evpi_inputs <- plot_evpi(evpi,
          decision_vars = "NPV_women_ISFM_income_inputs_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

men_evpi_inputs <- plot_evpi(evpi,
          decision_vars = "NPV_men_ISFM_income_inputs_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

women_evpi_knowledge <- plot_evpi(evpi,
          decision_vars = "NPV_women_ISFM_income_knowledge_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

men_evpi_knowledge <- plot_evpi(evpi,
          decision_vars = "NPV_men_ISFM_income_knowledge_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

women_evpi_labor <- plot_evpi(evpi,
          decision_vars = "NPV_women_ISFM_income_labor_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

men_evpi_labor <- plot_evpi(evpi,
          decision_vars = "NPV_men_ISFM_income_labor_based",  
          unit = "Ghana cedis",
          bar_color = "yellow4",
          base_size = 10)

women_evpi
men_evpi
women_evpi_land
men_evpi_land
women_evpi_inputs
men_evpi_inputs
women_evpi_knowledge
men_evpi_knowledge
women_evpi_labor
men_evpi_labor

```
In all cases, there were no variables with a positive EVPI, hence we do not need to plot them.


